---
title: "o-well Data set"
output:
  pdf_document:
    extra_dependencies:
    - float
    - flafter
    fig_caption: yes
    fig_width: 6
    fig_height: 6
  word_document: default
indent: yes
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancypagestyle{plain}{\pagestyle{fancy}}
- \headsep = 60pt
sections: yes
editor_options:
  chunk_output_type: inline
---
\fancyhead[RO,RE]{Justin Pickel \\ STATS 419 \\ Dr.Shaffer }
\fancyhead[LO,LE]{`r format(Sys.time(), '%B %d, %Y')`}
\fancyhead[CO,CE]{Assignment: o-well}
\setlength\parindent{24pt}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

my.source = 'github';
github.path = "https://raw.githubusercontent.com/jrpickel/WSU_STATS419_2021/";
table1 =  paste0(github.path,"main/dataset/o-well/pdf_num1.txt");
table2 = paste0(github.path,"main/dataset/o-well/pdf_num2.txt");
locations = paste0(github.path,"main/dataset/o-well/wells-location.txt");
```

```{r,read tables,include=FALSE}
owell.metals = read.table(table1)
owell.chemistry = read.table(table2,header=TRUE)
owell.location = read.table(locations,header=TRUE,sep = "|")
```

```{r,fig.cap="Wells metal data"}
head(owell.metals) # Table 1.
```

```{r,fig.cap="Wells chemistry data"}
head(owell.chemistry) # Table 2.
```

The following tables show the first five entries of the data frame. To have a deeper understand of the data I have made a data dictionary along with a brief summary of the data.  

# Summary of data
In March 2012 ground water samples were collected from twenty-three different wells along the coastal area of Gulf of Aqaba, Saudi Arabia. Most of the well are privately owned, dug in swallow aquifers and are located in relative close proximity tot eh east coast, except for well 23 which is a deeply dug well.
\begin{quote}
"Results of dissolved metals and physicochemical properties of groundwater samples are presented in Table
1 and Table 2. Metal contents in groundwater samples were low throughout the sampling wells and they are
within the range listed for waters suitable for drinking water (WHO, 2008). High concentrations of these metals
have been found in the adjacent soil samples and geologic units (Table 3 and Table 4). This suggests that the
primary source of dissolved metals to groundwater is not probably metals leached from the surrounding rocks
and soils, but rather released from aquifer materials (water-rock interaction). It may also suggest that
groundwater aquifer is not significantly recharged from surface runoff or the recharge rate from surface water is
low or negligible. This is consistent with the low and erratic annual precipitation rate occurred in the region." - (Journal of Applied Science and Agriculture,2013)
\end{quote}

# Data Dictionary
The following dictionary has been made to better understand the columns names of the two tables. Each value in the table is expressed in $\mu$g\textbackslash L \par
## Table 1 
As - Arsenic  
B - Boron  
Ba - Barium  
Be - Beryllium  
Cd - Cadmium  
Co - Cobalt  
Cr - Chromium  
Cu - Copper 
Fe - Iron  
Hg - Mercury  
Mn - Manganese  
Mo - Molybdenum  
Pb - Lead  
Se - selenium  
Zn - Zinc\par
## Table 2
well - The well number   
ph - The ph of the well water  
Eh - The redox of the well  
TDS - Total dissolved solids  
Ca - Calcium   
K - Potassium  
Mg - Magnesium  
Na - Sodium  
HCO3 - Bicarbonate  
Cl - Chlorine    
SO4 - Sulfate   
NO3 - Nitrate  
F - Fluorine 
Po4 - Phosphate  
TH - Total Hardness  
TA - Total Alkalinity  
TS - Total Solids  
SS - Suspended Solids  
COD - Chemical Oxygen Demand  
BOD - Biological Oxygen Demand  
DO - Dissolved Oxygen  

```{r,location data,include=FALSE}
location.data = owell.location[c(2,3)]
cbind(owell.metals,location.data)
cbind(owell.chemistry,location.data)
```

```{r,fault distances,include=FALSE}
owell.metals$fault = c(2189,1900,3629,4674,720,4895,2064,4494,868,5882,5813,5665,670,493,4818,1114,2705,5582,2539,4466 ,6009,3313,2697)
owell.chemistry$fault = c(2189,1900,3629,4674,720,4895,2064,4494,868,5882,5813,5665,670,493,4818,1114,2705,5582,2539,4466 ,6009,3313,2697)
```

```{r,rock formations,include=FALSE}
owell.metals$geology = c("Sand","Alkaline","Alkaline","Alkaline","Granite","Volcanic","Gypsum","Alkaline","Gypsum","Volcanic","Alkaline,Granite","Granite","Gypsum","Gypsum","Volcanic","Alkaline,Volcanic","Gypsum","Sand","Sand","Sand","Sand","Sand","Alkaline,Granite")

owell.chemistry$geology = c("Sand","Alkaline","Alkaline","Alkaline","Granite","Volcanic","Gypsum","Alkaline","Gypsum","Volcanic","Alkaline,Granite","Granite","Gypsum","Gypsum","Volcanic","Alkaline,Volcanic","Gypsum","Sand","Sand","Sand","Sand","Sand","Alkaline,Granite")
```

```{r,saving list as RDS,include=FALSE}
wells = list()

wells$saudi$chemistry = owell.chemistry
wells$saudi$chemistry = owell.metals

saveRDS(wells,"wells.rds")
```
