---
title: "r-intro"
output:
  pdf_document:
    extra_dependencies:
    - float
    - flafter
    fig_caption: yes
    fig_width: 6
    fig_height: 6
  word_document: default
indent: yes
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancypagestyle{plain}{\pagestyle{fancy}}
- \headsep = 60pt
sections: yes
editor_options:
  chunk_output_type: inline
---
\fancyhead[RO,RE]{Justin Pickel \\ WSU-ID: 11594028 \\ STATS 419 \\ Dr.Shaffer \\ `r format(Sys.time(), '%B %d, %Y')` }
\setlength\parindent{24pt}

\setlength\parindent{24pt}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
library(tinytex)
#install.packages("stringr")            # Install stringr package
#install.packages('igraph')
library(stringr)                        # Load stringr package
library(igraph)
library(devtools)

my.source = 'github';
github.path = "https://raw.githubusercontent.com/jrpickel/WSU_STATS419_2021/";
source_url( paste0(github.path,"main/functions/functions-intro.R") );
```
# Handshake
This function will determine how many handshakes are possible between a given number of people. The arguments for the function ``handshake(n,plotme=F)`` are ``n`` which is a positive integer and ``plotme`` which is a boolean value to determine if a graphical representation should be displayed the default value for plotme is ``plotme=FALSE``. The following formula is what the handshake function uses to determine the number of possible handshakes.
\[
\frac{n(n-1)}{2}
\]
```{r}
handshake(4,T)
```
